---
title: "Benchmark dose"
author: "En-Hsuan Lu"
date: "2023-07-07"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```

## dose response data
Ford et al. 2022 chemicals

```{r read}
datafolder <- "BBMD input"
set.seed(1234)
```

##2,4,5-Trichlorophenol

```{r 2,4,5-trichlorophenol}
#NOEL: 100 mg/kg/day
#Liver and kidney pathology
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.245tcp <- 100/AF_pod_noael_deter
cat("bmd.245tcp",quantile(bmd.245tcp,prob=c(0.5,0.05,0.95)),"\n")

#UFs=10
AF_pod_dur_sub <- rlnorm(10000,meanlog = log(2), sdlog=log(4^(1/1.645)))

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat, subchronic
#M: 0.18 kg; F: 0.124 kg
#average: 0.152 kg
#P50 = (70/0.152)^0.3 = 6.294719
#P95/P50 = (70/0.152)^0.04 = 1.277998
AF_interBS_sub.sdrat <- rlnorm(10000,meanlog = log(6.30), sdlog=log(1.28^(1/1.645)))

hed.245tcp <- 100/(AF_pod_noael_deter*AF_pod_dur_sub*AF_interBS_sub.sdrat)
cat("hed.245tcp",quantile(hed.245tcp,prob=c(0.5,0.05,0.95)),"\n")
```

##4,6-Dinitro-o-cresol
POD based on human

```{r 4,6-dinitro-o-cresol}
#LOAEL: 0.8 mg/kg/day
#AF LOAEL to NOAEL: 10
#Reduced body weight, excessive sweating, fatigue, decreased appetite, elevated BMR, and greenish-yellow coloration of the conjunctivae
AF_loael_noael <- rlnorm(10000, meanlog = log(10), sdlog = log(3))
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.645)))
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.dnoc.cont <- 0.8/(AF_loael_noael*AF_pod_noael_cont)
cat("bmd.dnoc.cont",quantile(bmd.dnoc.cont,prob=c(0.5,0.05,0.95)),"\n")
bmd.dnoc.deter <- 0.8/(AF_loael_noael*AF_pod_noael_deter)
cat("bmd.dnoc.deter",quantile(bmd.dnoc.deter,prob=c(0.5,0.05,0.95)),"\n")

#UFs=10
AF_pod_dur_sub <- rlnorm(10000,meanlog = log(2), sdlog=log(4^(1/1.645)))

#human
hed.dnoc.cont <- 0.8/(AF_loael_noael*AF_pod_dur_sub*AF_pod_noael_cont)
cat("hed.dnoc.cont",quantile(hed.dnoc.cont,prob=c(0.5,0.05,0.95)),"\n")
hed.dnoc.deter <- 0.8/(AF_loael_noael*AF_pod_dur_sub*AF_pod_noael_deter)
cat("hed.dnoc.deter",quantile(hed.dnoc.deter,prob=c(0.5,0.05,0.95)),"\n")
```

##ALDRIN
```{r aldrin}
### POD from BBMD
##nonneoplastic liver lesion
bmd.aldrin.df <- fread(file.path(datafolder,"aldrin-nonneoplastic-liver-lesion-bmds.csv"))
bmd.aldrin <- bmd.aldrin.df[15001:30000,]$model_average[sample.int(15000,10000)]
cat("bmd.aldrin",quantile(bmd.aldrin,prob=c(0.5,0.05,0.95)),"\n")

#AF: use WHO body size scaling
#human: 70 kg
#Rat (Osborne-Mendel strain): Male 0.514 kg; Female 0.389 kg
#Average bw: 0.4515 kg
#P50 = (70/0.4515)^0.3 = 4.541
#P95/P50 = (70/0.4515)^0.04 = 1.224
AF_interBS_omrat <- rlnorm(10000,meanlog = log(4.541), sdlog=log(1.224^(1/1.645)))

hed.aldrin <- bmd.aldrin/AF_interBS_omrat
cat("hed.aldrin",quantile(hed.aldrin,prob=c(0.5,0.05,0.95)),"\n")
```

##AZINPHOS-METHYL
male
```{r azinphos-methyl}
### POD from BBMD
##erythrocyte ChE
bmd.azin.df <- fread(file.path(datafolder,"azinphos-methyl-male-bmds.csv"))
bmd.azin <- bmd.azin.df$model_average[sample.int(15000,10000)]
cat("bmd.azin",quantile(bmd.azin,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg
#P50 = (70/10.8)^0.3 = 1.751872
#P95/P50 = (70/10.8)^0.04 = 1.077623
AF_interBS_dog.m <- rlnorm(10000,meanlog = log(1.75), sdlog=log(1.08^(1/1.645)))

hed.azin <- bmd.azin/AF_interBS_dog.m
cat("hed.azin",quantile(hed.azin,prob=c(0.5,0.05,0.95)),"\n")
```

##CHLORPYRIFOS
plasma AChE
male
```{r chlorpyrifos}
#plasma AChE
#male
bmd.chlo.plasma_m.df <- fread(file.path(datafolder,"chlopyrifos-plasma-male-bmds.csv"))
bmd.chlo.plasma_m <- bmd.chlo.plasma_m.df[1:14997,]$model_average[sample.int(14997,10000)]
cat("bmd.chlo.plasma_m",quantile(bmd.chlo.plasma_m,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sherman-strain rat: use Fischer 344 rat, male 0.38 kg
#P50 = (70/0.38)^0.3 = 4.781832
#P95/P50 = (70/0.38)^0.04 = 1.232005
AF_interBS_fischerrat.m <- rlnorm(10000,meanlog = log(4.78), sdlog=log(1.23^(1/1.645)))

hed.chlo.pla.m <- bmd.chlo.plasma_m/AF_interBS_fischerrat.m
cat("hed.chlo.pla.m",quantile(hed.chlo.pla.m,prob=c(0.5,0.05,0.95)),"\n")
```

##CHLORPYRIFOS
plasma AChE
female
```{r chlorpyrifos}
#plasma AChE
#female
bmd.chlo.plasma_f.df <- fread(file.path(datafolder,"chlopyrifos-plasma-female-bmds.csv"))
bmd.chlo.plasma_f <- bmd.chlo.plasma_f.df$model_average[sample.int(15000,10000)]
cat("bmd.chlo.plasma_f",quantile(bmd.chlo.plasma_f,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sherman-strain rat: use Fischer 344 rat, female 0.229 kg
#P50 = (70/0.229)^0.3 = 5.566456
#P95/P50 = (70/0.229)^0.04 = 1.257218
AF_interBS_fischerrat.f <- rlnorm(10000,meanlog = log(5.57), sdlog=log(1.26^(1/1.645)))

hed.chlo.pla.f <- bmd.chlo.plasma_f/AF_interBS_fischerrat.f
cat("hed.chlo.pla.f",quantile(hed.chlo.pla.f,prob=c(0.5,0.05,0.95)),"\n")
```

##CHLORPYRIFOS
RBC AChE
male
```{r chlorpyrifos}
###the most sensitive POD/BMDL
#RBC AChE
#male
bmd.chlo.rbc_m.df <- fread(file.path(datafolder,"chlopyrifos-rbc-male-bmds.csv"))
bmd.chlo.rbc_m <- bmd.chlo.rbc_m.df$model_average[sample.int(15000,10000)]
cat("bmd.chlo.rbc_m",quantile(bmd.chlo.rbc_m,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sherman-strain rat: use Fischer 344 rat, male 0.38 kg
#P50 = (70/0.38)^0.3 = 4.781832
#P95/P50 = (70/0.38)^0.04 = 1.232005
AF_interBS_fischerrat.m <- rlnorm(10000,meanlog = log(4.78), sdlog=log(1.23^(1/1.645)))

hed.chlo.rbc.m <- bmd.chlo.rbc_m/AF_interBS_fischerrat.m
cat("hed.chlo.rbc.m",quantile(hed.chlo.rbc.m,prob=c(0.5,0.05,0.95)),"\n")
```

##CHLORPYRIFOS
RBC AChE
female
```{r chlorpyrifos}
#RBC AChE
#female
bmd.chlo.rbc_f.df <- fread(file.path(datafolder,"chlopyrifos-rbc-female-bmds.csv"))
bmd.chlo.rbc_f <- bmd.chlo.rbc_f.df$model_average[sample.int(15000,10000)]
cat("bmd.chlo.rbc_f",quantile(bmd.chlo.rbc_f,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sherman-strain rat: use Fischer 344 rat, female 0.229 kg
#P50 = (70/0.229)^0.3 = 5.566456
#P95/P50 = (70/0.229)^0.04 = 1.257218
AF_interBS_fischerrat.f <- rlnorm(10000,meanlog = log(5.57), sdlog=log(1.26^(1/1.645)))

hed.chlo.rbc.f <- bmd.chlo.rbc_f/AF_interBS_fischerrat.f
cat("hed.chlo.rbc.f",quantile(hed.chlo.rbc.f,prob=c(0.5,0.05,0.95)),"\n")
```

##CHLORPYRIFOS
brain AChE
male
```{r chlorpyrifos}
#brain AChE
#male
bmd.chlo.brain_m.df <- fread(file.path(datafolder,"chlopyrifos-brain-male-bmds.csv"))
bmd.chlo.brain_m <- bmd.chlo.brain_m.df$model_average[sample.int(15000,10000)]
cat("bmd.chlo.brain_m",quantile(bmd.chlo.brain_m,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sherman-strain rat: use Fischer 344 rat, male 0.38 kg
#P50 = (70/0.38)^0.3 = 4.781832
#P95/P50 = (70/0.38)^0.04 = 1.232005
AF_interBS_fischerrat.m <- rlnorm(10000,meanlog = log(4.78), sdlog=log(1.23^(1/1.645)))

hed.chlo.br.m <- bmd.chlo.brain_m/AF_interBS_fischerrat.m
cat("hed.chlo.br.m",quantile(hed.chlo.br.m,prob=c(0.5,0.05,0.95)),"\n")
```

##CHLORPYRIFOS
brain AChE
female
```{r chlorpyrifos}
#brain AChE
#female
bmd.chlo.brain_f.df <- fread(file.path(datafolder,"chlopyrifos-brain-female-bmds.csv"))
bmd.chlo.brain_f <- bmd.chlo.brain_f.df$model_average[sample.int(15000,10000)]
cat("bmd.chlo.brain_f",quantile(bmd.chlo.brain_f,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sherman-strain rat: use Fischer 344 rat, female 0.229 kg
#P50 = (70/0.229)^0.3 = 5.566456
#P95/P50 = (70/0.229)^0.04 = 1.257218
AF_interBS_fischerrat.f <- rlnorm(10000,meanlog = log(5.57), sdlog=log(1.26^(1/1.645)))

hed.chlo.br.f <- bmd.chlo.brain_f/AF_interBS_fischerrat.f
cat("hed.chlo.br.f",quantile(hed.chlo.br.f,prob=c(0.5,0.05,0.95)),"\n")
```

##DIAZINON

```{r diazinon}
#NOAEL: 0.065 mg/kg/day
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.63)))

bmd.diaz <- 0.065/AF_pod_noael_cont
cat("bmd.diaz",quantile(bmd.diaz,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat
#M: 0.523 kg; F: 0.338 kg
#average: 0.4305 kg
#P50 = (70/0.4305)^0.3 = 4.606144
#P95/P50 = (70/0.4305)^0.04 = 1.226
AF_interBS_sdrat <- rlnorm(10000,meanlog = log(4.61), sdlog=log(1.226^(1/1.645)))

hed.diaz <- 0.065/(AF_pod_noael_cont*AF_interBS_sdrat)
cat("hed.diaz",quantile(hed.diaz,prob=c(0.5,0.05,0.95)),"\n")
```

##Dibutyl Phthalate

```{r DBP}
#NOAEL: 125 mg/kg/day
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.dbp <- 125/AF_pod_noael_deter
cat("bmd.dbp",quantile(bmd.dbp,prob=c(0.5,0.05,0.95)),"\n")

#UFs=3
AF_pod_dur_sub <- rlnorm(10000,meanlog = log(2), sdlog=log(4^(1/1.645)))

#AF: WHO body size scaling
#human: 70 kg
#subchronic Male SD rat: 0.267 kg
#P50 = (70/0.267)^0.3 = 5.315891
#P95/P50 = (70/0.267)^0.04 = 1.249521
AF_interBS_sub.sdrat.m <- rlnorm(10000,meanlog = log(5.32), sdlog=log(1.25^(1/1.645)))

hed.dbp <- 125/(AF_pod_noael_deter*AF_pod_dur_sub*AF_interBS_sub.sdrat.m)
cat("hed.dbp",quantile(hed.dbp,prob=c(0.5,0.05,0.95)),"\n")
```

##DICOFOL
male
```{r dicofol}
#male
bmd.dicofol_m.df <- fread(file.path(datafolder,"dicofol-male-bmds.csv"))
bmd.dicofol_m <- bmd.dicofol_m.df[1:14998,]$model_average[sample.int(14998,10000)]
cat("bmd.dicofol_m",quantile(bmd.dicofol_m,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg
#P50 = (70/10.8)^0.3 = 1.751872
#P95/P50 = (70/10.8)^0.04 = 1.077623
AF_interBS_dog.m <- rlnorm(10000,meanlog = log(1.75), sdlog=log(1.08^(1/1.645)))

hed.dicofol.m <- bmd.dicofol_m/AF_interBS_dog.m
cat("hed.dicofol.m",quantile(hed.dicofol.m,prob=c(0.5,0.05,0.95)),"\n")
```

##DICOFOL
female
```{r dicofol}
###the most sensitive POD/BMDL
#female
bmd.dicofol_f.df <- fread(file.path(datafolder,"dicofol-female-bmds.csv"))
bmd.dicofol_f <- bmd.dicofol_f.df[1:14993,]$model_average[sample.int(14993,10000)]
cat("bmd.dicofol_f",quantile(bmd.dicofol_f,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: female 10.1 kg
#P50 = (70/10.1)^0.3 = 1.787446
#P95/P50 = (70/10.1)^0.04 = 1.080516
AF_interBS_dog.f <- rlnorm(10000,meanlog = log(1.79), sdlog=log(1.08^(1/1.645)))

hed.dicofol.f <- bmd.dicofol_f/AF_interBS_dog.f
cat("hed.dicofol.f",quantile(hed.dicofol.f,prob=c(0.5,0.05,0.95)),"\n")
```

##DIELDRIN
liver bw ratio
```{r dieldrin}
### POD from BBMD
##liver bw ratio
bmd.diel.lir.df <- fread(file.path(datafolder,"dieldrin-liver-bw-ratio-bmds.csv"))
bmd.diel.lir <- bmd.diel.lir.df$model_average[sample.int(15000,10000)]
cat("bmd.diel.lir",quantile(bmd.diel.lir,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#rat: use Fischer 344 rat, male 0.38 kg; female 0.229 kg
#average animal bw: 0.3045 kg
#P50 = (70/0.3045)^0.3 = 5.11
#P95/P50 = (70/0.3045)^0.04 = 1.243
AF_interBS_fischerrat <- rlnorm(10000,meanlog = log(5.11), sdlog=log(1.243^(1/1.645)))

hed.diel.lir <- bmd.diel.lir/AF_interBS_fischerrat
cat("hed.diel.lir",quantile(hed.diel.lir,prob=c(0.5,0.05,0.95)),"\n")
```

##DIELDRIN
liver weight
```{r dieldrin}
##liver weight
bmd.diel.liw.df <- fread(file.path(datafolder,"dieldrin-liver-weight-bmds.csv"))
bmd.diel.liw <- bmd.diel.liw.df$model_average[sample.int(15000,10000)]
cat("bmd.diel.liw",quantile(bmd.diel.liw,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#rat: use Fischer 344 rat, male 0.38 kg; female 0.229 kg
#average animal bw: 0.3045 kg
#P50 = (70/0.3045)^0.3 = 5.11
#P95/P50 = (70/0.3045)^0.04 = 1.243
AF_interBS_fischerrat <- rlnorm(10000,meanlog = log(5.11), sdlog=log(1.243^(1/1.645)))

hed.diel.liw <- bmd.diel.liw/AF_interBS_fischerrat
cat("hed.diel.liw",quantile(hed.diel.liw,prob=c(0.5,0.05,0.95)),"\n")
```

##Disulfoton

```{r disulfoton}
#LEL: 0.04 mg/kg/day
#ChE inhibition, optic nerve degeneration
#both
#AF LOAEL to NOAEL: 10
AF_loael_noael <- rlnorm(10000, meanlog = log(10), sdlog = log(3))
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.645)))
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.disul.cont <- 0.04/(AF_loael_noael*AF_pod_noael_cont)
cat("bmd.disul.cont",quantile(bmd.disul.cont,prob=c(0.5,0.05,0.95)),"\n")
bmd.disul.deter <- 0.04/(AF_loael_noael*AF_pod_noael_cont)
cat("bmd.disul.deter",quantile(bmd.disul.deter,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#rat: use Fischer 344 rat, male 0.38 kg; female 0.229 kg
#average animal bw: 0.3045 kg
#P50 = (70/0.3045)^0.3 = 5.11
#P95/P50 = (70/0.3045)^0.04 = 1.243
AF_interBS_fischerrat <- rlnorm(10000,meanlog = log(5.11), sdlog=log(1.243^(1/1.645)))

hed.disul.cont <- 0.04/(AF_loael_noael*AF_pod_noael_cont*AF_interBS_fischerrat)
cat("hed.disul.cont",quantile(hed.disul.cont,prob=c(0.5,0.05,0.95)),"\n")
hed.disul.deter <- 0.04/(AF_loael_noael*AF_pod_noael_deter*AF_interBS_fischerrat)
cat("hed.disul.deter",quantile(hed.disul.deter,prob=c(0.5,0.05,0.95)),"\n")
```

##Endosulfan
glomerulonephrosis
male
```{r endosulfan}
#glomerulonephrosis
#male
bmd.endo.glo_m.df <- fread(file.path(datafolder,"endosulfan-glomerulo-male-bmds.csv"))
bmd.endo.glo_m <- bmd.endo.glo_m.df[15001:29931,]$model_average[sample.int(14931,10000)]
cat("bmd.endo.glo_m",quantile(bmd.endo.glo_m,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat: M: 0.523 kg
#P50 = (70/0.523)^0.3 = 4.344891
#P95/P50 = (70/0.523)^0.04 = 1.216365
AF_interBS_sdrat.m <- rlnorm(10000,meanlog = log(4.35), sdlog=log(1.22^(1/1.645)))

hed.endo.glo.m <- bmd.endo.glo_m/AF_interBS_sdrat.m
cat("hed.endo.glo.m",quantile(hed.endo.glo.m,prob=c(0.5,0.05,0.95)),"\n")
```

##Endosulfan
reduced bw
male
```{r endosulfan}
###the most sensitive POD/BMDL
#reduced bw
#male
bmd.endo.bw_m.df <- fread(file.path(datafolder,"endosulfan-bw-male-bmds.csv"))
bmd.endo.bw_m <- bmd.endo.bw_m.df[1:14999,]$model_average[sample.int(14999,10000)]
cat("bmd.endo.bw_m",quantile(bmd.endo.bw_m,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat: M: 0.523 kg
#P50 = (70/0.523)^0.3 = 4.344891
#P95/P50 = (70/0.523)^0.04 = 1.216365
AF_interBS_sdrat.m <- rlnorm(10000,meanlog = log(4.35), sdlog=log(1.22^(1/1.645)))

hed.endo.bw.m <- bmd.endo.bw_m/AF_interBS_sdrat.m
cat("hed.endo.bw.m",quantile(hed.endo.bw.m,prob=c(0.5,0.05,0.95)),"\n")
```

##Endosulfan
reduced bw
female
```{r endosulfan}
#female
bmd.endo.bw_f.df <- fread(file.path(datafolder,"endosulfan-bw-female-bmds.csv"))
bmd.endo.bw_f <- bmd.endo.bw_f.df[1:14997,]$model_average[sample.int(14997,10000)]
cat("bmd.endo.bw_f",quantile(bmd.endo.bw_f,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat: F: 0.338 kg
#P50 = (70/0.338)^0.3 = 4.952841
#P95/P50 = (70/0.338)^0.04 = 1.237791
AF_interBS_sdrat.f <- rlnorm(10000,meanlog = log(4.95), sdlog=log(1.24^(1/1.645)))

hed.endo.bw.f <- bmd.endo.bw_f/AF_interBS_sdrat.f
cat("hed.endo.bw.f",quantile(hed.endo.bw.f,prob=c(0.5,0.05,0.95)),"\n")
```

##Endrin

```{r endrin}
#NOEL: 0.025 mg/kg/day
#Mild histological lesions in liver, occasional convulsions
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.endrin <- 0.025/AF_pod_noael_deter
cat("bmd.endrin",quantile(bmd.endrin,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#dog: use beagle dog, male 10.8 kg; female 10.1 kg 
#average animal bw: 10.45 kg
#P50 = (70/10.45)^0.3 = 1.769
#P95/P50 = (70/10.45)^0.04 = 1.079
AF_interBS_dog <- rlnorm(10000,meanlog = log(1.769), sdlog=log(1.079^(1/1.645)))

hed.endrin <- 0.025/(AF_pod_noael_deter*AF_interBS_dog)
cat("hed.endrin",quantile(hed.endrin,prob=c(0.5,0.05,0.95)),"\n")
```

##Ethion
POD based on human

```{r ethion}
#NOEL: 0.05 mg/kg/day
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.63)))

bmd.ethion <- 0.05/AF_pod_noael_cont
cat("bmd.ethion",quantile(bmd.ethion,prob=c(0.5,0.05,0.95)),"\n")

#human
hed.ethion <- 0.05/AF_pod_noael_cont
cat("hed.ethion",quantile(hed.ethion,prob=c(0.5,0.05,0.95)),"\n")
```

##Heptachlor
male
```{r heptachlor}
##HEPTACHLOR
### POD from BBMD
##IgM response in male
bmd.heptachlor.df <- fread(file.path(datafolder,"heptachlor-igm-response-bmds.csv"))
bmd.heptachlor <- bmd.heptachlor.df$model_average[sample.int(14910,10000)]
cat("bmd.heptachlor",quantile(bmd.heptachlor,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Subchronic female SD rat: F: 0.204 kg
#P50 = (70/0.204)^0.3 = 5.763
#P95/P50 = (70/0.204)^0.04 = 1.263
AF_interBS_sub.sdrat.f <- rlnorm(10000,meanlog = log(5.763), sdlog=log(1.263^(1/1.645)))

hed.heptachlor <- bmd.heptachlor/AF_interBS_sub.sdrat.f
cat("hed.heptachlor",quantile(hed.heptachlor,prob=c(0.5,0.05,0.95)),"\n")
```

##Heptachlor epoxide

```{r heptachlorepoxide}
#LEL: 0.0125 mg/kg/day
#AF LOAEL to NOAEL: 10
AF_loael_noael <- rlnorm(10000, meanlog = log(10), sdlog = log(3))
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.645)))

bmd.hepepo <- 0.0125/(AF_loael_noael*AF_pod_noael_cont)
cat("bmd.hepepo",quantile(bmd.hepepo,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg; female 10.1 kg
#average animal bw: 10.45 kg
#P50 = (70/10.45)^0.3 = 1.769
#P95/P50 = (70/10.45)^0.04 = 1.079
AF_interBS_dog <- rlnorm(10000,meanlog = log(1.769), sdlog=log(1.079^(1/1.645)))

hed.hepepo <- 0.0125/(AF_loael_noael*AF_pod_noael_cont*AF_interBS_dog)
cat("hed.hepepo",quantile(hed.hepepo,prob=c(0.5,0.05,0.95)),"\n")
```

##Methoxychlor

```{r methoxychlor}
###litter size loss
bmd.mtc.df <- fread(file.path(datafolder,"methoxychlor-bmds.csv"))
bmd.mtc <- bmd.mtc.df[15001:30000,]$model_average[sample.int(15000,10000)]
cat("bmd.mtc",quantile(bmd.mtc, prob=c(0.5,0.05,0.95)))

#AF: WHO body size scaling
#human: 70 kg
#Female New Zealand white rabbit: 3.93 kg
#P50 = (70/3.93)^0.3 = 2.373
#P95/P50 = (70/3.93)^0.04 = 1.122
AF_interBS_rabbit.f <- rlnorm(10000,meanlog = log(2.373), sdlog=log(1.122^(1/1.645)))

hed.mtc <- bmd.mtc/AF_interBS_rabbit.f
cat("hed.mtc",quantile(hed.mtc,prob=c(0.5,0.05,0.95)),"\n")
```

##P,P'-DDT

```{r DDT}
#NOEL: 0.05 mg/kg/day
#Increasing hepatocellular hypertrophy, especially centrilobularly, increased cytoplasmic oxyphilia, and peripheral basophilic cytoplasmic granules (based on H and E paraffin sections)
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.ddt <- 0.05/AF_pod_noael_deter
cat("bmd.ddt",quantile(bmd.ddt,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat, subchronic: M: 0.18 kg; F: 0.124 kg
#average: 0.152 kg
#P50 = (70/0.152)^0.3 = 6.294719
#P95/P50 = (70/0.152)^0.04 = 1.277998
AF_interBS_sub.sdrat <- rlnorm(10000,meanlog = log(6.30), sdlog=log(1.28^(1/1.645)))

hed.ddt <- 0.05/(AF_pod_noael_deter*AF_interBS_sub.sdrat)
cat("hed.ddt",quantile(hed.ddt,prob=c(0.5,0.05,0.95)),"\n")
```

##Parathion
POD based on human

```{r parathion}
#NOAEL: 0.064 mg/kg/day
#Cholinesterase / Decreased cholinesterase activity
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.63)))

bmd.parathion <- 0.064/AF_pod_noael_cont
cat("bmd.parathion",quantile(bmd.parathion,prob=c(0.5,0.05,0.95)),"\n")

#human
hed.parathion <- 0.064/AF_pod_noael_cont
cat("hed.parathion",quantile(hed.parathion,prob=c(0.5,0.05,0.95)),"\n")
```

##Pentachlorophenol
chronic inflammation
male
```{r pentachlorophenol}
### POD from BBMD
##chronic inflammation
bmd_m.pcp.infla.df <- fread(file.path(datafolder,"pentachlorophenol-chronic-inflammation-male-bmds.csv"))
bmd_m.pcp.infla <- bmd_m.pcp.infla.df[15001:30000,]$model_average[sample.int(15000,10000)]
cat("bmd_m.pcp.infla",quantile(bmd_m.pcp.infla,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg
#P50 = (70/10.8)^0.3 = 1.751872
#P95/P50 = (70/10.8)^0.04 = 1.077623
AF_interBS_dog.m <- rlnorm(10000,meanlog = log(1.75), sdlog=log(1.08^(1/1.645)))

hed.pcp.infla.m <- bmd_m.pcp.infla/AF_interBS_dog.m
cat("hed.pcp.infla.m",quantile(hed.pcp.infla.m,prob=c(0.5,0.05,0.95)),"\n")
```

##Pentachlorophenol
chronic inflammation
female
```{r pentachlorophenol}
bmd_f.pcp.infla.df <- fread(file.path(datafolder,"pentachlorophenol-chronic-inflammation-female-bmds.csv"))
bmd_f.pcp.infla <- bmd_f.pcp.infla.df[15001:29984,]$model_average[sample.int(14984,10000)] #14984
cat("bmd_f.pcp.infla",quantile(bmd_f.pcp.infla,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: female 10.1 kg
#P50 = (70/10.1)^0.3 = 1.787446
#P95/P50 = (70/10.1)^0.04 = 1.080516
AF_interBS_dog.f <- rlnorm(10000,meanlog = log(1.79), sdlog=log(1.08^(1/1.645)))

hed.pcp.infla.f <- bmd_f.pcp.infla/AF_interBS_dog.f
cat("hed.pcp.infla.f",quantile(hed.pcp.infla.f,prob=c(0.5,0.05,0.95)),"\n")
```

##Pentachlorophenol
cyto vacuolization
male
```{r pentachlorophenol}
bmd_m.pcp.cyto.df <- fread(file.path(datafolder,"pentachlorophenol-cyto-vacuolization-male-bmds.csv"))
bmd_m.pcp.cyto <- (bmd_m.pcp.cyto.df[15001:29995,]$model_average)[sample.int(14995,10000)] #14995
cat("bmd_m.pcp.cyto",quantile(bmd_m.pcp.cyto,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg
#P50 = (70/10.8)^0.3 = 1.751872
#P95/P50 = (70/10.8)^0.04 = 1.077623
AF_interBS_dog.m <- rlnorm(10000,meanlog = log(1.75), sdlog=log(1.08^(1/1.645)))

hed.pcp.cyto.m <- bmd_m.pcp.cyto/AF_interBS_dog.m
cat("hed.pcp.cyto.m",quantile(hed.pcp.cyto.m,prob=c(0.5,0.05,0.95)),"\n")
```

##Pentachlorophenol
cyto vacuolization
female
```{r pentachlorophenol}
bmd_f.pcp.cyto.df <- fread(file.path(datafolder,"pentachlorophenol-cyto-vacuolization-female-bmds.csv"))
bmd_f.pcp.cyto <- (bmd_f.pcp.cyto.df[15001:29964,]$model_average)[sample.int(14964,10000)] #14964
cat("bmd_f.pcp.cyto",quantile(bmd_f.pcp.cyto,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: female 10.1 kg
#P50 = (70/10.1)^0.3 = 1.787446
#P95/P50 = (70/10.1)^0.04 = 1.080516
AF_interBS_dog.f <- rlnorm(10000,meanlog = log(1.79), sdlog=log(1.08^(1/1.645)))

hed.pcp.cyto.f <- bmd_f.pcp.cyto/AF_interBS_dog.f
cat("hed.pcp.cyto.f",quantile(hed.pcp.cyto.f,prob=c(0.5,0.05,0.95)),"\n")
```

##Pentachlorophenol
liver pigment
male
``````{r pentachlorophenol}
bmd_m.pcp.liver.df <- fread(file.path(datafolder,"pentachlorophenol-liver-pigment-male-bmds.csv"))
bmd_m.pcp.liver <- (bmd_m.pcp.liver.df[15001:30000,]$model_average)[sample.int(15000,10000)]
cat("bmd_m.pcp.liver",quantile(bmd_m.pcp.liver,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg
#P50 = (70/10.8)^0.3 = 1.751872
#P95/P50 = (70/10.8)^0.04 = 1.077623
AF_interBS_dog.m <- rlnorm(10000,meanlog = log(1.75), sdlog=log(1.08^(1/1.645)))

hed.pcp.liver.m <- bmd_m.pcp.liver/AF_interBS_dog.m
cat("hed.pcp.liver.m",quantile(hed.pcp.liver.m,prob=c(0.5,0.05,0.95)),"\n")
```

##Pentachlorophenol
liver pigment
female
``````{r pentachlorophenol}
bmd_f.pcp.liver.df <- fread(file.path(datafolder,"pentachlorophenol-liver-pigment-female-bmds.csv"))
bmd_f.pcp.liver <- (bmd_f.pcp.liver.df[15001:30000,]$model_average)[sample.int(15000,10000)]
cat("bmd_f.pcp.liver",quantile(bmd_f.pcp.liver,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: female 10.1 kg
#P50 = (70/10.1)^0.3 = 1.787446
#P95/P50 = (70/10.1)^0.04 = 1.080516
AF_interBS_dog.f <- rlnorm(10000,meanlog = log(1.79), sdlog=log(1.08^(1/1.645)))

hed.pcp.liver.f <- bmd_f.pcp.liver/AF_interBS_dog.f
cat("hed.pcp.liver.f",quantile(hed.pcp.liver.f,prob=c(0.5,0.05,0.95)),"\n")
```

#combine BMD samples in a dataframe
19 chemicals
35 endpoints

```{r dataframe}
bmd.samp.df <- data.frame(bmd.245tcp, 
                          bmd.dnoc.cont, bmd.dnoc.deter, 
                          bmd.aldrin, bmd.azin, 
                          bmd.chlo.plasma_m, bmd.chlo.plasma_f, bmd.chlo.rbc_m, bmd.chlo.rbc_f, bmd.chlo.brain_m, bmd.chlo.brain_f, 
                          bmd.diaz, bmd.dbp, 
                          bmd.dicofol_m, bmd.dicofol_f, 
                          bmd.diel.lir, bmd.diel.liw, 
                          bmd.disul.cont, bmd.disul.deter, 
                          bmd.endo.glo_m, bmd.endo.bw_m, bmd.endo.bw_f, 
                          bmd.endrin, bmd.ethion, bmd.heptachlor, bmd.hepepo, bmd.mtc,
                          bmd.ddt, bmd.parathion, 
                          bmd_m.pcp.infla, bmd_f.pcp.infla, bmd_m.pcp.cyto, bmd_f.pcp.cyto, bmd_m.pcp.liver, bmd_f.pcp.liver)

#add CAS and make columns to rows
bmd.samp.df.t <- data.frame(t(bmd.samp.df))

cas.df <- c("95-95-4", 
            "534-52-1", "534-52-1",
            "309-00-2", "86-50-0",
            "2921-88-2", "2921-88-2", "2921-88-2", "2921-88-2", "2921-88-2", "2921-88-2", 
            "333-41-5", "84-74-2",
            "115-32-2", "115-32-2",
            "60-57-1", "60-57-1", 
            "298-04-4", "298-04-4",
            "115-29-7", "115-29-7", "115-29-7", 
            "72-20-8", "563-12-2", "76-44-8", "1024-57-3", "72-43-5",
            "50-29-3", "56-38-2", 
            "87-86-5", "87-86-5", "87-86-5", "87-86-5", "87-86-5", "87-86-5")
end.df <- c("2,4,5-Trichlorophenol",
            "4,6-Dinitro-o-cresol continous endpoint","4,6-Dinitro-o-cresol dichotomous endpoint",
            "Aldrin","Azinphos-methyl",
            "Chlorpyrifos male plasma AChE","Chlorpyrifos female plasma AChE","Chlorpyrifos male RBC AChE","Chlorpyrifos female RBC AChE","Chlorpyrifos male brain AChE","Chlorpyrifos female brain AChE",
            "Diazinon","Dibutyl phthalate",
            "Dicofol male","Dicofol female",
            "Dieldrin increased liver-to-bw ratio","Dieldrin increased liver weight",
            "Disulfoton continous endpoint","Disulfoton dichotomous endpoint",
            "Endosulfan male glomerulonephrosis","Endosulfan male reduced bw","Endosulfan female reduced bw",
            "Endrin","Ethion","Heptachlor","Heptachlor epoxide","Methoxychlor",
            "p,p'-DDT","Parathion",
            "Pentachlorophenol male chronic inflammation","Pentachlorophenol female chronic inflammation","Pentachlorophenol male cytoplasmic vacuolization","Pentachlorophenol female cytoplasmic vacuolization","Pentachlorophenol male liver pigment","Pentachlorophenol female liver pigment")

bmd.samp.t.cas <- cbind(bmd.samp.df.t, cas.df, end.df)
colnames(bmd.samp.t.cas)[10001:10002] <- c("CAS","endpoint")

write.csv(bmd.samp.t.cas, file="bmd samples CAS by row.csv", row.names = FALSE)
```

#quantile samples

```{r quantile}
bmd.quan <- as.data.frame(t(apply(bmd.samp.df, 2, quantile, probs=c(0.5, 0.05, 0.95))))
bmd.quan.df <- cbind(bmd.quan, end.df, cas.df)
colnames(bmd.quan.df)[4:5] <- c("endpoint","CAS")
write.csv(bmd.quan.df, file="BMD quantile.csv", row.names = FALSE)

bbmd.quan <- bmd.quan.df[-c(1:3,12,13,18,19,23,24,26,28,29),]
write.csv(bbmd.quan, file="BBMD quantile.csv", row.names=FALSE)
```

#combine HED samples in a dataframe

```{r dataframe}
hed.samp.df <- data.frame(hed.245tcp, 
                          hed.dnoc.cont, hed.dnoc.deter, 
                          hed.aldrin, hed.azin, 
                          hed.chlo.pla.m, hed.chlo.pla.f, hed.chlo.rbc.m, hed.chlo.rbc.f, hed.chlo.br.m, hed.chlo.br.f, 
                          hed.diaz, hed.dbp, 
                          hed.dicofol.m, hed.dicofol.f,
                          hed.diel.lir, hed.diel.liw, 
                          hed.disul.cont, hed.disul.deter,
                          hed.endo.glo.m, hed.endo.bw.m, hed.endo.bw.f,
                          hed.endrin, hed.ethion, 
                          hed.heptachlor, hed.hepepo, hed.mtc, 
                          hed.ddt, hed.parathion, 
                          hed.pcp.infla.m, hed.pcp.infla.f, hed.pcp.cyto.m, hed.pcp.cyto.f, hed.pcp.liver.m, hed.pcp.liver.f)

#add CAS and make columns to rows
hed.samp.df.t <- data.frame(t(hed.samp.df))
hed.samp.t.cas <- cbind(hed.samp.df.t, cas.df, end.df)
colnames(hed.samp.t.cas)[10001:10002] <- c("CAS","endpoint")

write.csv(hed.samp.t.cas, file="hed samples CAS by row.csv", row.names = FALSE)
```