---
title: "Benchmark dose"
author: "En-Hsuan Lu"
date: "2023-07-17"
output: html_document
---

##set seed

```{r read}
set.seed(1234)
```

##2,4,5-Trichlorophenol

```{r 2,4,5-trichlorophenol}
#NOEL: 100 mg/kg/day
#Liver and kidney pathology
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.245tcp <- 100/AF_pod_noael_deter
cat("bmd.245tcp",quantile(bmd.245tcp,prob=c(0.5,0.05,0.95)),"\n")

AF_pod_dur_sub <- rlnorm(10000,meanlog = log(2), sdlog=log(4^(1/1.645)))

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat, subchronic
#M: 0.18 kg; F: 0.124 kg
#average: 0.152 kg
#P50 = (70/0.152)^0.3 = 6.294719
#P95/P50 = (70/0.152)^0.04 = 1.277998
AF_interBS_sub.sdrat <- rlnorm(10000,meanlog = log(6.30), sdlog=log(1.28^(1/1.645)))

hed.245tcp <- 100/(AF_pod_noael_deter*AF_pod_dur_sub*AF_interBS_sub.sdrat)
cat("hed.245tcp",quantile(hed.245tcp,prob=c(0.5,0.05,0.95)),"\n")
```

##4,6-Dinitro-o-cresol
POD based on human

```{r 4,6-dinitro-o-cresol}
#LOAEL: 0.8 mg/kg/day
#AF LOAEL to NOAEL: 10
#Reduced body weight, excessive sweating, fatigue, decreased appetite, elevated BMR, and greenish-yellow coloration of the conjunctivae
AF_loael_noael <- rlnorm(10000, meanlog = log(10), sdlog = log(3)^(1/1.645))
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.645)))
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

AF_pod_dur_sub <- rlnorm(10000,meanlog = log(2), sdlog=log(4^(1/1.645)))

bmd.dnoc.cont <- 0.8/(AF_loael_noael*AF_pod_noael_cont)
cat("bmd.dnoc.cont",quantile(bmd.dnoc.cont,prob=c(0.5,0.05,0.95)),"\n")
bmd.dnoc.deter <- 0.8/(AF_loael_noael*AF_pod_noael_deter)
cat("bmd.dnoc.deter",quantile(bmd.dnoc.deter,prob=c(0.5,0.05,0.95)),"\n")

#human
hed.dnoc.cont <- 0.8/(AF_loael_noael*AF_pod_dur_sub*AF_pod_noael_cont)
cat("hed.dnoc.cont",quantile(hed.dnoc.cont,prob=c(0.5,0.05,0.95)),"\n")
hed.dnoc.deter <- 0.8/(AF_loael_noael*AF_pod_dur_sub*AF_pod_noael_deter)
cat("hed.dnoc.deter",quantile(hed.dnoc.deter,prob=c(0.5,0.05,0.95)),"\n")
```

##ALDRIN
```{r aldrin}
#LOAEL: 0.025
#LOAEL to NOAEL: 10
AF_loael_noael <- rlnorm(10000, meanlog = log(10), sdlog = log(3)^(1/1.645))
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.aldrin <- 0.025/(AF_loael_noael*AF_pod_noael_deter)
cat("bmd.aldrin",quantile(bmd.aldrin,prob=c(0.5,0.05,0.95)),"\n")

#AF: use WHO body size scaling
#human: 70 kg
#Rat (Osborne-Mendel strain): Male 0.514 kg; Female 0.389 kg
#Average bw: 0.4515 kg
#P50 = (70/0.4515)^0.3 = 4.541
#P95/P50 = (70/0.4515)^0.04 = 1.224
AF_interBS_omrat <- rlnorm(10000,meanlog = log(4.541), sdlog=log(1.224^(1/1.645)))

hed.aldrin <- 0.025/(AF_loael_noael*AF_pod_noael_deter*AF_interBS_omrat)
cat("hed.aldrin",quantile(hed.aldrin,prob=c(0.5,0.05,0.95)),"\n")
```

##AZINPHOS-METHYL
male
```{r azinphos-methyl}
#BMD: 0.48 mg/kg/day
#BMDL:0.3 mg/kg/day
#P50=BMD=0.48
#P95/P50=BMD/BMDL=0.48/0.3=1.6
bmd.azin <- rlnorm(10000,meanlog = log(0.48), sdlog=log(1.6)^(1/1.645))
cat("bmd.azin",quantile(bmd.azin,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg
#P50 = (70/10.8)^0.3 = 1.751872
#P95/P50 = (70/10.8)^0.04 = 1.077623
AF_interBS_dog.m <- rlnorm(10000,meanlog = log(1.75), sdlog=log(1.08^(1/1.645)))

hed.azin <- bmd.azin/AF_interBS_dog.m
cat("hed.azin",quantile(hed.azin,prob=c(0.5,0.05,0.95)),"\n")
```

##CHLORPYRIFOS

```{r chlorpyrifos}
#NOAEL: 0.1
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.chlo <- 0.1/AF_pod_noael_deter
cat("bmd.chlo",quantile(bmd.chlo,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sherman-strain rat: use Fischer 344 rat, male 0.38 kg, female 0.229 kg
#average bw: 0.3045
#P50 = (70/0.3045)^0.3 = 5.11038
#P95/P50 = (70/0.3045)^0.04 = 1.242969
AF_interBS_fischerrat <- rlnorm(10000,meanlog = log(5.11), sdlog=log(1.24^(1/1.645)))

hed.chlo <- 0.1/(AF_pod_noael_deter*AF_interBS_fischerrat)
cat("hed.chlo",quantile(hed.chlo,prob=c(0.5,0.05,0.95)),"\n")
```

##DIAZINON

```{r diazinon}
#NOAEL: 0.065 mg/kg/day
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.63)))

bmd.diaz <- 0.065/AF_pod_noael_cont
cat("bmd.diaz",quantile(bmd.diaz,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat
#M: 0.523 kg; F: 0.338 kg
#average: 0.4305 kg
#P50 = (70/0.4305)^0.3 = 4.606144
#P95/P50 = (70/0.4305)^0.04 = 1.226
AF_interBS_sdrat <- rlnorm(10000,meanlog = log(4.61), sdlog=log(1.226^(1/1.645)))

hed.diaz <- 0.065/(AF_pod_noael_cont*AF_interBS_sdrat)
cat("hed.diaz",quantile(hed.diaz,prob=c(0.5,0.05,0.95)),"\n")
```

##Dibutyl Phthalate

```{r DBP}
#NOAEL: 125 mg/kg/day
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.dbp <- 125/AF_pod_noael_deter
cat("bmd.dbp",quantile(bmd.dbp,prob=c(0.5,0.05,0.95)),"\n")

AF_pod_dur_sub <- rlnorm(10000,meanlog = log(2), sdlog=log(4^(1/1.645)))

#AF: WHO body size scaling
#human: 70 kg
#subchronic Male SD rat: 0.267 kg
#P50 = (70/0.267)^0.3 = 5.315891
#P95/P50 = (70/0.267)^0.04 = 1.249521
AF_interBS_sub.sdrat.m <- rlnorm(10000,meanlog = log(5.32), sdlog=log(1.25^(1/1.645)))

hed.dbp <- 125/(AF_pod_noael_deter*AF_pod_dur_sub*AF_interBS_sub.sdrat.m)
cat("hed.dbp",quantile(hed.dbp,prob=c(0.5,0.05,0.95)),"\n")
```

##DICOFOL

```{r dicofol}
#NOAEL: 0.12
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.dicofol <- 0.12/AF_pod_noael_deter
cat("bmd.dicofol",quantile(bmd.dicofol,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg, female 10.1 kg
#average bw: 10.45
#P50 = (70/10.45)^0.3 = 1.769272
#P95/P50 = (70/10.45)^0.04 = 1.079044
AF_interBS_dog <- rlnorm(10000,meanlog = log(1.77), sdlog=log(1.08^(1/1.645)))

hed.dicofol <- 0.12/(AF_pod_noael_deter*AF_interBS_dog)
cat("hed.dicofol",quantile(hed.dicofol,prob=c(0.5,0.05,0.95)),"\n")
```

##DIELDRIN

```{r dieldrin}
#NOAEL: 0.005
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.63)))

bmd.diel <- 0.005/AF_pod_noael_cont
cat("bmd.diel",quantile(bmd.diel,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#rat: use Fischer 344 rat, male 0.38 kg; female 0.229 kg
#average animal bw: 0.3045 kg
#P50 = (70/0.3045)^0.3 = 5.11
#P95/P50 = (70/0.3045)^0.04 = 1.243
AF_interBS_fischerrat <- rlnorm(10000,meanlog = log(5.11), sdlog=log(1.243^(1/1.645)))

hed.diel <- 0.005/(AF_pod_noael_cont*AF_interBS_fischerrat)
cat("hed.diel",quantile(hed.diel,prob=c(0.5,0.05,0.95)),"\n")
```

##Disulfoton

```{r disulfoton}
#LEL: 0.04 mg/kg/day
#ChE inhibition, optic nerve degeneration
#both
#AF LOAEL to NOAEL: 10
AF_loael_noael <- rlnorm(10000, meanlog = log(10), sdlog = log(3)^(1/1.645))
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.645)))
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.disul.cont <- 0.04/(AF_loael_noael*AF_pod_noael_cont)
cat("bmd.disul.cont",quantile(bmd.disul.cont,prob=c(0.5,0.05,0.95)),"\n")
bmd.disul.deter <- 0.04/(AF_loael_noael*AF_pod_noael_cont)
cat("bmd.disul.deter",quantile(bmd.disul.deter,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#rat: use Fischer 344 rat, male 0.38 kg; female 0.229 kg
#average animal bw: 0.3045 kg
#P50 = (70/0.3045)^0.3 = 5.11
#P95/P50 = (70/0.3045)^0.04 = 1.243
AF_interBS_fischerrat <- rlnorm(10000,meanlog = log(5.11), sdlog=log(1.243^(1/1.645)))

hed.disul.cont <- 0.04/(AF_loael_noael*AF_pod_noael_cont*AF_interBS_fischerrat)
cat("hed.disul.cont",quantile(hed.disul.cont,prob=c(0.5,0.05,0.95)),"\n")
hed.disul.deter <- 0.04/(AF_loael_noael*AF_pod_noael_deter*AF_interBS_fischerrat)
cat("hed.disul.deter",quantile(hed.disul.deter,prob=c(0.5,0.05,0.95)),"\n")
```

##Endosulfan

```{r endosulfan}
#NOAEL: 0.6
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.645)))
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.endo.cont <- 0.6/AF_pod_noael_cont
cat("bmd.endo.cont",quantile(bmd.endo.cont,prob=c(0.5,0.05,0.95)),"\n")
bmd.endo.deter <- 0.6/AF_pod_noael_cont
cat("bmd.endo.deter",quantile(bmd.endo.deter,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat: M: 0.523 kg, F: 0.338 kg
#average bw: 0.4305
#P50 = (70/0.4305)^0.3 = 4.606144
#P95/P50 = (70/0.4305)^0.04 = 1.225872
AF_interBS_sdrat <- rlnorm(10000,meanlog = log(4.61), sdlog=log(1.23^(1/1.645)))

hed.endo.cont <- 0.6/(AF_pod_noael_cont*AF_interBS_sdrat)
cat("hed.endo.cont",quantile(hed.endo.cont,prob=c(0.5,0.05,0.95)),"\n")
hed.endo.deter <- 0.6/(AF_pod_noael_cont*AF_interBS_sdrat)
cat("hed.endo.deter",quantile(hed.endo.deter,prob=c(0.5,0.05,0.95)),"\n")
```

##Endrin

```{r endrin}
#NOEL: 0.025 mg/kg/day
#Mild histological lesions in liver, occasional convulsions
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.endrin <- 0.025/AF_pod_noael_deter
cat("bmd.endrin",quantile(bmd.endrin,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#dog: use beagle dog, male 10.8 kg; female 10.1 kg 
#average animal bw: 10.45 kg
#P50 = (70/10.45)^0.3 = 1.769
#P95/P50 = (70/10.45)^0.04 = 1.079
AF_interBS_dog <- rlnorm(10000,meanlog = log(1.769), sdlog=log(1.079^(1/1.645)))

hed.endrin <- 0.025/(AF_pod_noael_deter*AF_interBS_dog)
cat("hed.endrin",quantile(hed.endrin,prob=c(0.5,0.05,0.95)),"\n")
```

##Ethion
POD based on human

```{r ethion}
#NOEL: 0.05 mg/kg/day
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.63)))

bmd.ethion <- 0.05/AF_pod_noael_cont
cat("bmd.ethion",quantile(bmd.ethion,prob=c(0.5,0.05,0.95)),"\n")

#human
hed.ethion <- 0.05/AF_pod_noael_cont
cat("hed.ethion",quantile(hed.ethion,prob=c(0.5,0.05,0.95)),"\n")
```

##Heptachlor
female
```{r heptachlor}
#LOAEL: 0.03
#AF LOAEL to NOAEL: 3
AF_loael_noael_3 <- rlnorm(10000, meanlog = log(3), sdlog = log(3)^(1/1.645))
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.645)))

bmd.heptachlor <- 0.03/(AF_loael_noael_3*AF_pod_noael_cont)
cat("bmd.heptachlor",quantile(bmd.heptachlor,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Subchronic female SD rat: F: 0.204 kg
#P50 = (70/0.204)^0.3 = 5.763
#P95/P50 = (70/0.204)^0.04 = 1.263
AF_interBS_sub.sdrat.f <- rlnorm(10000,meanlog = log(5.763), sdlog=log(1.263^(1/1.645)))

hed.heptachlor <- 0.03/(AF_loael_noael_3*AF_pod_noael_cont*AF_interBS_sub.sdrat.f)
cat("hed.heptachlor",quantile(hed.heptachlor,prob=c(0.5,0.05,0.95)),"\n")
```

##Heptachlor epoxide

```{r heptachlorepoxide}
#LEL: 0.0125 mg/kg/day
#AF LOAEL to NOAEL: 10
AF_loael_noael <- rlnorm(10000, meanlog = log(10), sdlog = log(3)^(1/1.645))
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.645)))

bmd.hepepo <- 0.0125/(AF_loael_noael*AF_pod_noael_cont)
cat("bmd.hepepo",quantile(bmd.hepepo,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg; female 10.1 kg
#average animal bw: 10.45 kg
#P50 = (70/10.45)^0.3 = 1.769
#P95/P50 = (70/10.45)^0.04 = 1.079
AF_interBS_dog <- rlnorm(10000,meanlog = log(1.769), sdlog=log(1.079^(1/1.645)))

hed.hepepo <- 0.0125/(AF_loael_noael*AF_pod_noael_cont*AF_interBS_dog)
cat("hed.hepepo",quantile(hed.hepepo,prob=c(0.5,0.05,0.95)),"\n")
```

##Methoxychlor

```{r methoxychlor}
#NOEL: 5.01
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.mtc <- 5.01/AF_pod_noael_deter
cat("bmd.mtc",quantile(bmd.mtc, prob=c(0.5,0.05,0.95)))

#AF: WHO body size scaling
#human: 70 kg
#Female New Zealand white rabbit: 3.93 kg
#P50 = (70/3.93)^0.3 = 2.373
#P95/P50 = (70/3.93)^0.04 = 1.122
AF_interBS_rabbit.f <- rlnorm(10000,meanlog = log(2.373), sdlog=log(1.122^(1/1.645)))

hed.mtc <- 5.01/(AF_pod_noael_deter*AF_interBS_rabbit.f)
cat("hed.mtc",quantile(hed.mtc,prob=c(0.5,0.05,0.95)),"\n")
```

##P,P'-DDT
```{r DDT}
#NOEL: 0.05 mg/kg/day
#Increasing hepatocellular hypertrophy, especially centrilobularly, increased cytoplasmic oxyphilia, and peripheral basophilic cytoplasmic granules (based on H and E paraffin sections)
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.ddt <- 0.05/AF_pod_noael_deter
cat("bmd.ddt",quantile(bmd.ddt,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#Sprague-Dawley rat, subchronic: M: 0.18 kg; F: 0.124 kg
#average: 0.152 kg
#P50 = (70/0.152)^0.3 = 6.294719
#P95/P50 = (70/0.152)^0.04 = 1.277998
AF_interBS_sub.sdrat <- rlnorm(10000,meanlog = log(6.30), sdlog=log(1.28^(1/1.645)))

hed.ddt <- 0.05/(AF_pod_noael_deter*AF_interBS_sub.sdrat)
cat("hed.ddt",quantile(hed.ddt,prob=c(0.5,0.05,0.95)),"\n")
```

##Parathion
POD based on human

```{r parathion}
#NOAEL: 0.064 mg/kg/day
#Cholinesterase / Decreased cholinesterase activity
AF_pod_noael_cont <- rlnorm(10000,meanlog = log(1/3), sdlog=log(4.7^(1/1.63)))

bmd.parathion <- 0.064/AF_pod_noael_cont
cat("bmd.parathion",quantile(bmd.parathion,prob=c(0.5,0.05,0.95)),"\n")

#human
hed.parathion <- 0.064/AF_pod_noael_cont
cat("hed.parathion",quantile(hed.parathion,prob=c(0.5,0.05,0.95)),"\n")
```

##Pentachlorophenol

```{r pentachlorophenol}
#LOAEL: 1.5
#AF LOAEL to NOAEL: 3
AF_loael_noael_3 <- rlnorm(10000, meanlog = log(3), sdlog = log(3)^(1/1.645))
AF_pod_noael_deter <- rlnorm(10000,meanlog = log(2/9), sdlog=log(5^(1/1.645)))

bmd.pcp <- 1.5/(AF_loael_noael_3*AF_pod_noael_deter)
cat("bmd.pcp",quantile(bmd.pcp,prob=c(0.5,0.05,0.95)),"\n")

#AF: WHO body size scaling
#human: 70 kg
#beagle dog: male 10.8 kg, female 10.1 kg
#average bw: 10.45
#P50 = (70/10.45)^0.3 = 1.769272
#P95/P50 = (70/10.45)^0.04 = 1.079044
AF_interBS_dog <- rlnorm(10000,meanlog = log(1.77), sdlog=log(1.08^(1/1.645)))

hed.pcp <- 1.5/(AF_loael_noael_3*AF_pod_noael_deter*AF_interBS_dog)
cat("hed.pcp",quantile(hed.pcp,prob=c(0.5,0.05,0.95)),"\n")
```

#combine BMD samples in a dataframe

```{r dataframe}
who_bmd.samp.df <- data.frame(bmd.245tcp, 
                              bmd.dnoc.cont, bmd.dnoc.deter, 
                              bmd.aldrin, bmd.azin, 
                              bmd.chlo, bmd.diaz, bmd.dbp, bmd.dicofol, bmd.diel, 
                              bmd.disul.cont, bmd.disul.deter,
                              bmd.endo.cont, bmd.endo.deter, 
                              bmd.endrin, bmd.ethion, 
                              bmd.heptachlor, bmd.hepepo, bmd.mtc,
                              bmd.ddt, bmd.parathion,
                              bmd.pcp)

#add CAS and make columns to rows
who_bmd.samp.df.t <- data.frame(t(who_bmd.samp.df))

cas.df <- c("95-95-4",
            "534-52-1","534-52-1",
            "309-00-2","86-50-0",
            "2921-88-2","333-41-5","84-74-2","115-32-2","60-57-1",
            "298-04-4","298-04-4",
            "115-29-7","115-29-7",
            "72-20-8","563-12-2","76-44-8","1024-57-3","72-43-5",
            "50-29-3","56-38-2",
            "87-86-5")
end.df <- c("2,4,5-Trichlorophenol",
            "4,6-Dinitro-o-cresol continuous endpoint","4,6-Dinitro-o-cresol dichotomous endpoint",
            "Aldrin","Azinphos-methyl",
            "Chlorpyrifos","Diazinon","Dibutyl phthalate","Dicofol","Dieldrin",
            "Disulfoton continous endpoint","Disulfoton dichotomous endpoint",
            "Endosulfan continous endpoint","Endosulfan dichotomous endpoint",
            "Endrin","Ethion","Heptachlor","Heptachlor epoxide","Methoxychlor","p,p'-DDT","Parathion","Pentachlorophenol")

who_bmd.samp.t.cas <- cbind(who_bmd.samp.df.t, cas.df, end.df)
colnames(who_bmd.samp.t.cas)[10001:10002] <- c("CAS","endpoint")

write.csv(who_bmd.samp.t.cas, file="WHO bmd samples CAS by row.csv", row.names = FALSE)
```

#BMD quantile

```{r quantile}
who.bmd.quan <- as.data.frame(t(apply(who_bmd.samp.df, 2, quantile, probs=c(0.5, 0.05, 0.95))))
who.bmd.quan.df <- cbind(who.bmd.quan, end.df, cas.df)
colnames(who.bmd.quan.df)[4:5] <- c("endpoint","CAS")
write.csv(who.bmd.quan.df, file="WHO BMD quantile.csv", row.names = FALSE)
```

#combine HED samples in a dataframe

```{r dataframe}
who_hed.samp.df <- data.frame(hed.245tcp, 
                              hed.dnoc.cont, hed.dnoc.deter, 
                              hed.aldrin, hed.azin, 
                              hed.chlo, hed.diaz, hed.dbp, hed.dicofol, hed.diel, 
                              hed.disul.cont, hed.disul.deter,
                              hed.endo.cont, hed.endo.deter, 
                              hed.endrin, hed.ethion, 
                              hed.heptachlor, hed.hepepo, hed.mtc,
                              hed.ddt, hed.parathion,
                              hed.pcp)

#add CAS and make columns to rows
who_hed.samp.df.t <- data.frame(t(who_hed.samp.df))
who_hed.samp.t.cas <- cbind(who_hed.samp.df.t, cas.df, end.df)
colnames(who_hed.samp.t.cas)[10001:10002] <- c("CAS","endpoint")

write.csv(who_hed.samp.t.cas, file="WHO hed samples CAS by row.csv", row.names = FALSE)
```